Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-07-31 mu/article/article.tex
Making english corrections

\begin{abstract}
We present a theoretical study of spin-velocity injection of a pure
spin current induced by  a linearly
polarized beam of light that impinges normally at the surface of  two 50\%
hydrogenated noncentrosymmetric 2D graphene 
structures, labeled Up and Alt.
The hydrogenation opens an energy gap in both structures.
We analyze two possibilities, one in which by fixing the spin along a
chosen direction, the resulting spin-velocity is calculated, 
and the other one, where we chose the spin-velocity direction
on the surface plane, and calculate its speed and
spin orientation. We do so by changing the energy $\hbar\go$ 
and polarization angle $\ga$  of the incoming  beam. 
The results are calculated in a full
electronic band structure scheme within the DFT in the LDA approximation.
The maxima of the spin-velocity speed is obtained when
$\hbar\go=0.084$\,eV and $\ga=35^\circ$ for Up,
and
$\hbar\go=0.720$\,eV and $\ga=150^\circ$ for Alt.
We find a speed of
668\,Km/s
and 
645\,Km/s
for Up and Alt, respectively, when the spin points perpendicular to
the surface. Also,
the response is maximized
by fixing the spin-velocity direction along a high symmetry Cartesian axis,
obtaining a speed of 
688\,Km/s 
and the spin  pointing at 13$^\circ$ form the surface normal, for Up,
and
906\,Km/s 
and the spin pointing at 60$^\circ$ from the surface normal, for Alt.
These speed values are larger than those of bulk semiconductors, like
CdSe and GaAs, thus 
making the hydrogenated graphene structures
 excellent candidates for spintronics
applications.
\end{abstract}

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state 
media plays the determining role adding a new degree of freedom to the
conventional charge manipulation.\cite{wolfSC04,fabianAPS07}
% 
At present there is an increasing interest in attain the same level of control
over the transport of spin at micro or nano scales as has been done for the
flow of charge in typical electronic devices.\cite{awschalomNP2007} Some
semiconductor spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08} and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC).
% 
One of the difficulties to achieve the development of spin current and PSC
semiconductor devices is the fact that the spin relaxation time in a
semiconducting media is short disabling the spin transport and then resulting
in a no observable spin current.\cite{murakamiSc03}
% 
In PSCs there is no net
motion of charge; spin-up electrons move in a given direction while spin-down
electrons travel in the opposite one. 
This effect can result from one photon absorption of linearly polarized light
by a semiconductor, with filled valence bands and empty conduction bands,
illuminated by light with photon energy larger than the energy gap.
For instance,
this phenomena can result from spin
injection,\cite{malPRB03} Hall Effects,\cite{sinovaPRB04} interference of two
optical beams,\cite{bhatPRL00, najmaiePRB03} or one photon absorption of
linearly polarized light\cite{bhatPRL05} and has been observed in gallium
arsenide (GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

Graphene, an allotrope of carbon with hexagonal 2D lattice structure presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08} being then a
perfect platform to be used in two-dimensions electronic systems; however most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that the band gap of graphene can be opened by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals,\cite{varykhalovPRB10} and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
% 
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
% 
In this paper we present two 50\% hydrogenated graphene noncentrosymmetric
structures both presenting a discernible band gap: the Up structure,
shown in Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer
only in the upper side of the structure while the Alt structure, shown
in Fig. \ref{fig:alt-struc}, has hydrogen alternating in the upper and bottom
sides of the carbon slab.\cite{zapataPSB2016}

@@

Using those structures we address a theoretical study of the 
spin velocity
injection (SVI) 
by one-photon absorption of linearly polarized light.
Because we have 2D structures we made the analysis for tow cases. The first is
fixing the spin of the electrons along the $z$ Cartesian direction with the
velocity directed on the surface of the structure in the $xy$ plane. The second
is fixing the spin velocity in the $x$ or $y$ direction and the spin directed
in $xyz$.
% 
The SVI is an optical effect that quantifies the velocity at which a PSC moves
along the Cartesian direction $\mathrm{a}$ with the spin of electron polarized
along the Cartesian direction $\mathrm{b}$. One photon absorption of linearly
polarized light can promote an even distribution of electrons in $\mathbf{k}$ space
regardless the symmetry of the material resulting in a 
null electrical current.\cite{bhatPRL05}
Then, the electrons excited to the conduction bands at opposite $\mathbf{k}$
points will result in opposite spin polarizations producing no net spin
injection.\cite{bhatPRL05} If the crystalline structure of the material is
noncentrosymmetric the spin polarization injected at a given $\mathbf{k}$ point
not necessarily 
 vanishes,\cite{alvaradoPRL85, schmiedeskampPRL88} 
and then, since the velocities of electrons at opposite $\mathbf{k}$ points are
opposite, a PSC will be produced. Since the structures presented
here are noncentrosymmetric, they are good candidates in which this effect can
be induced.

@@

$\hat \bfr$ the position operator,  $\hat H_0$ the unperturbed
ground state Hamiltonian, and

@@

of the spin-splited conduction bands inherent to the 

@@

are quasidegenerate  states and we take $\ge\to 0$ at the end of the

@@

The matriz elements of any operator $\calo$ are given by

@@

with $\hbar \omega_{n}(\mathbf{k})$ the energy of the

@@

above expression, one can show that

@@

where repeated cartesian are summed, and 

@@

K^{\mathrm{ab}}_{nm}(\mathbf{-k}) = K^{\mathrm{ab*}}_{nm}(\mathbf{k}). 

@@

that follows from time-reversal invariance. The prime $'$ in the sum means that
$c$ and $c'$ are quasi degenerate states and the sum only covers these states.
Since $\mu^{\mathrm{abcd}}(\omega)$ is real we have that

@@

 is the same as Eq. (3) of Bhat et al.\cite{bhatPRL05}

@@

%\subsection{Spin velocity injection} % (fold)
%\label{sec:theory-pure_spin_current}
Now, we define the spin velocity injection (SVI) as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
that gives the velocity, along the direction $\mathrm{a}$, at which the spin
moves polarized along the direction $\mathrm{b}$. The carrier injection rate
$\dot n(\go)$ is written as,\cite{nastosPRB05}
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega),
\label{eq:dotn}
\end{equation}
where the tensor 
\begin{equation}\label{eq:xi}
\begin{aligned}
\xi^{\mathrm{ab}}(\omega)
&
=
\frac{2\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}}
 \sum_{vc}
r^{\mathrm{a}}_{vc'}(\bfk)  
r^{\mathrm{b}}_{cv }(\bfk)  
\delta(\omega-\omega_{cv}(\bfk) 
, 
\end{aligned}
\end{equation}
is related to the imaginary part of the linear optical 
response tensor by
$\mathrm{Im}[\ge^{\rma\rmb}(\go)]=2\pi\ge_0\hbar\xi^{\rma\rmb}(\go)$.

The function $\calv^{\rma\rmb}(\go)$ allow us to quantify two very
important aspects of PSC. On one hand, we can fix the spin direction
along $\mathrm{b}$,   
and calculate the resulting electron velocity.  
On the other hand, we can fix de velocity of the electron
along $\mathrm{b}$,  and study the resulting direction along which the
spin is polarized.
To this end, the added advantage of  2D structures, besides choosing
them noncentrosymmetric, is that we can use an incoming linearly
polarized beam of light
at normal incidence, and use the  direction of the polarized  electric
field to control $\calv^{\rma\rmb}(\go)$.
Indeed, writing 
$\bfE(\omega) = E_0(\omega)(\cos\ga\,\hat\bfx+\sin\ga\,\hat\bfy)$
where $\alpha$ is the polarization angle, we obtain from
 Eq. \eqref{eq:vab-w}
that
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= 
\frac{2}{\hbar\xi(\go)}
\left(\mu^{\mathrm{abxx}}(\omega)\cos^{2}\alpha + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}\alpha + 
\mu^{\mathrm{abxy}}(\omega)\sin 2\alpha\right)
\label{eq:vab-aw}
\end{align}
\end{widetext}
as for the structures chosen in this article,
$\xi^{\mathrm{xx}}(\go)=\xi^{\mathrm{yy}}(\go)\equiv\xi(\go)$, and $\xi^{\mathrm{xy}}(\go)=0$.
Now, we formalize our two options for $\calv^{\rma\rmb}(\go)$.

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define the magnitude of
the electron velocity in plane
with the spin polarized along the $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\gs^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
% where this angle is measured in the counter-clockwise direction from the
% positive $x$. 
We also define two special angles
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular
the incoming polarization, 
respectively. The subscript $\gs^\rmb$ denotes the spin along $\rmb$.

@@

Fixing the calculated velocity along $\rma=x$ or $\rma=y$
 we define it's corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented
in the $xyz$ system's coordinates
according to a polar angle
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
= 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{align}
and an azimuthal angle
\begin{align}
\varphi_{\mathrm{a}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}

@@

We preset the results of $\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ and
$\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ for the C$_{16}$H$_{8}$-up  and
C$_{16}$H$_{8}$-alt structures being both noncentrosymmetric semi-infinite 2D
carbon systems with 50\% hydrogenation in different arrangements. We recall
that the Up structure has hydrogen atoms only on the upper side of the carbon
sheet while the Alt structure has alternating hydrogen atoms on the upper and
bottom sides. Also we take the hexagonal carbon lattice to be on the $xy$ plane
for both structures, and the carbon-hydrogen bonds on the perpendicular $xz$
plane, as depicted in Figs. \ref{fig:up-struc} and \ref{fig:alt-struc}. The
coordinates for the Up and Alt unit cells of the structures are presented in
Tables \ref{tab:up-unitcell} and \ref{tab:alt-unitcell}.

We calculated the self-consistent ground state and the Kohn-Sham states using
density functional theory in the local density approximation (DFT-LDA) with a
planewave basis using the ABINIT code \cite{gonzeCPC09}.
% 
We used Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space
Gaussian pseudopotentials \cite{hartwigsenPRB98} including the spin-orbit
interaction needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented
in Eq. \eqref{eq:mu}.
% 
The convergence parameters for the calculations of our results corresponding to
the Up and Alt structures are cutoff energies of 40\,Ha and 65\,Ha, resulting
in LDA energy band gaps of 0.084\,eV and 0.718\,eV, respectively. The energy
eigenvalues and matrix elements for the Up and Alt structures were calculated
using 12802 $\mathbf{k}$ points and 14452 $\mathbf{k}$ points in the IBZ to
integrate $\mu^{\mathrm{abcd}}(\go)$ and $\xi^{\rma\rmb}(\go)$ using the
linearized analytic tetrahedron method (LATM).\cite{nastosPRB07} We neglect the
anomalous velocity term $\hbar(\bfgs\times\nabla V)/4m^2c^2$, where $V$ is the
crystal potential, in $\hat\bfv$ of Eq.~\eqref{z.1}, as this term is known to
give small contribution to PSC.\cite{bhatPRL05} Therefore,
$[\hat\bfv,\hat\bfS]=0$ and Eq.~\eqref{z.1} reduces to $\hat K^{\rma\rmb}=\hat
v^\rma \hat S^\rmb=\hat S^\rmb \hat v^\rma$. Finally, the prime in the sum of
Eq.~\eqref{eq:mu} is restricted to quasidegenerated conduction bands $c$ and
$c'$ that are closer than 30 meV, where this value is both a typical laser-
pulse energy width and the room-temperature energy.\cite{nastosPRB07}
% 
% It is known that the DFT with LDA approximation misestimates the energy band
% gap of semiconductors. To correct this one has to include the many-body
% interaction using i.e. the so-called GW approximation but this technique has a
% very high computational cost and is out of the scope in this paper.
% Nevertheless, DFT still remains a mainstream tool for calculating diverse
% properties derived from the electronic band structure.

@@

In this subsection we calculate $\calv_{\gs^\rmz}(\omega,\alpha)$, Eq.
\eqref{eq:vs-mag}, for the case where the spin is fixed along $z$, i.e.
directed perpendicularly to the surface of the Up and Alt structures. Also, we
calculate $\gamma_{\gs^\rmz}(\go,\ga)$, Eq. \eqref{eq:gamma-ang}, that
determines the direction along which the injected electrons move along the
surface of  each structure. We mention that we have also made the analysis for
the cases when the spin  is directed along $\mathrm{x}$ or $\mathrm{y}$,
finding similar qualitative results to those presented bellow.
%

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{upplots/up-vx-vy-w1}
\caption{For the Up structure we show
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$
(left scale, black line),  
$\theta_{\mathrm{a}} (\omega,\alpha)$
(right scale, red line), and
$\varphi_{\mathrm{a}} (\omega,\alpha)$,
(right scale, blue line),
vs. $\hbar\go$, for $\ga=35^\circ$, 
and $\rma=x,y$.
}
\label{fig:up-vab-comp-rtp-1}
\end{figure}

>>>>

\begin{abstract}
We present a theoretical study of spin-velocity injection of a pure
spin current induced by  a linearly
polarized beam of light that impinges normally on the surface of  two 50\%
hydrogenated noncentrosymmetric 2D graphene 
structures, labeled Up and Alt.
The hydrogenation opens an energy gap in both structures.
We analyze two possibilities: one in which by fixing the spin along a
chosen direction, the resulting spin-velocity is calculated, 
and another where we choose the spin-velocity direction
on the surface plane, and calculate its speed and
spin orientation. We do so by changing the energy $\hbar\go$ 
and polarization angle $\ga$  of the incoming  beam. 
The results are calculated within a full
electronic band structure scheme within the DFT in the LDA approximation.
The maxima of the spin-velocity speed are obtained when
$\hbar\go=0.084$\,eV and $\ga=35^\circ$ for Up,
and
$\hbar\go=0.720$\,eV and $\ga=150^\circ$ for Alt.
We find a speed of
668\,Km/s
and 
645\,Km/s
for Up and Alt, respectively, when the spin points perpendicularly to
the surface. Also,
the response is maximized
by fixing the spin-velocity direction along a high symmetry Cartesian axis,
obtaining a speed of 
688\,Km/s 
with the spin  pointing at 13$^\circ$ from the surface normal, for Up,
and
906\,Km/s 
and the spin pointing at 60$^\circ$ from the surface normal, for Alt.
These speed values are larger than those of bulk semiconductors, like
CdSe and GaAs, thus 
making the hydrogenated graphene structures
 excellent candidates for spintronics
applications.
\end{abstract}

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state 
media plays the determining role, adding a new degree of freedom to
conventional charge manipulation.\cite{wolfSC04,fabianAPS07}
% 
At present, there is an increasing interest in attaining the same level of
control over the transport of spin at micro or nano scales, as it has been done for
the flow of charge in typical electronic devices.\cite{awschalomNP2007} Some
semiconductor spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC).
% 
One of the difficulties to achieve the development of spin current and PSC
semiconductor devices is the fact that the spin relaxation time in a
semiconducting media is short, disabling the spin transport, and then resulting
in a non-observable spin current.\cite{murakamiSc03}
% 
In PSCs, there is no net
motion of charge; spin-up electrons move in a given direction, while spin-down
electrons travel in the opposite one. 
This effect can result from one photon absorption of linearly polarized light
by a semiconductor, with filled valence bands and empty conduction bands,
illuminated by light with photon energy larger than the energy gap.
This phenomenon can result from spin
injection,\cite{malPRB03} Hall Effects,\cite{sinovaPRB04} interference of two
optical beams,\cite{bhatPRL00, najmaiePRB03} or one photon absorption of
linearly polarized light\cite{bhatPRL05} and has been observed in gallium
arsenide (GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

Graphene, an allotrope of carbon with hexagonal 2D lattice structure, presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08}, being then a
perfect platform to be used in two-dimension electronic systems; however, most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that the band gap of graphene can be opened by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals,\cite{varykhalovPRB10} and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
% 
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
% 
In this paper, we present two 50\% hydrogenated graphene noncentrosymmetric
structures, both presenting a discernible band gap: the Up structure,
shown in Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer
only on the upper side of the structure, while the Alt structure, shown
in Fig. \ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom
sides of the carbon slab.\cite{zapataPSB2016}

@@

Using those structures, we address a theoretical study of the 
spin velocity
injection (SVI) 
by one-photon absorption of linearly polarized light.
Since we have 2D structures, we do the analysis for two cases. The first is by
fixing the spin of the electrons along the $z$ Cartesian direction, with the
velocity directed on the surface of the structure on the $xy$ plane. The second
is by fixing the spin velocity in the $x$ or $y$ direction and the spin directed
to $xyz$.
% 
The SVI is an optical effect that quantifies the velocity at which a PSC moves
along the Cartesian direction, $\mathrm{a}$, with the spin of electron polarized
along the Cartesian direction $\mathrm{b}$. One photon absorption of linearly
polarized light can promote an even distribution of electrons in $\mathbf{k}$ space,
regardless of the symmetry of the material, resulting in a 
null electrical current.\cite{bhatPRL05}
Then, the electrons excited to the conduction bands at opposite $\mathbf{k}$
points will result in opposite spin polarizations producing no net spin
injection.\cite{bhatPRL05} If the crystalline structure of the material is
noncentrosymmetric, the spin polarization injected at a given $\mathbf{k}$ point
not necessarily 
 vanishes,\cite{alvaradoPRL85, schmiedeskampPRL88} 
and then, since the velocities of electrons at opposite $\mathbf{k}$ points are
opposite, a PSC will be produced. Since the structures presented
here are noncentrosymmetric, they are good candidates in which this effect can
be induced.

@@

$\hat \bfr$ being the position operator and $\hat H_0$ the unperturbed
ground state Hamiltonian;

@@

of the spin-split conduction bands inherent to the 

@@

are quasidegenerate  states, and we take $\ge\to 0$ at the end of the

@@

The matrix elements of any operator $\calo$ are given by

@@

with $\hbar \omega_{n}(\mathbf{k})$ being the energy of the

@@

the above expression, one can show that

@@

where repeated Cartesians are summed, and 

@@

K^{\mathrm{ab}}_{nm}(\mathbf{-k}) = K^{\mathrm{ab*}}_{nm}(\mathbf{k}) 

@@

which follows from time-reversal invariance. The prime symbol $'$ in the sum means that
$c$ and $c'$ are quasi degenerate states, and the sum only covers these states.
Since $\mu^{\mathrm{abcd}}(\omega)$ is real, we have that

@@

is the same as Eq. (3) of Bhat et al.\cite{bhatPRL05},

@@

%\subsection{Spin velocity injection} % (fold)
%\label{sec:theory-pure_spin_current}
Now, we define the spin velocity injection (SVI) as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
which gives the velocity, along direction $\mathrm{a}$, at which the spin
moves in a polarized state along direction $\mathrm{b}$. The carrier injection rate
$\dot n(\go)$ is written as,\cite{nastosPRB05}
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega)
\label{eq:dotn}
\end{equation}
where the tensor 
\begin{equation}\label{eq:xi}
\begin{aligned}
\xi^{\mathrm{ab}}(\omega)
&
=
\frac{2\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}}
 \sum_{vc}
r^{\mathrm{a}}_{vc'}(\bfk)  
r^{\mathrm{b}}_{cv }(\bfk)  
\delta(\omega-\omega_{cv}(\bfk) 
, 
\end{aligned}
\end{equation}
is related to the imaginary part of the linear optical 
response tensor by
$\mathrm{Im}[\ge^{\rma\rmb}(\go)]=2\pi\ge_0\hbar\xi^{\rma\rmb}(\go)$.

The function $\calv^{\rma\rmb}(\go)$ allows us to quantify two very
important aspects of PSC. On one hand, we can fix the spin direction
along $\mathrm{b}$  
and calculate the resulting electron velocity.  
On the other hand, we can fix the velocity of the electron
along $\mathrm{b}$ and study the resulting direction along which the
spin is polarized.
To this end, the added advantage of  2D structures, besides choosing
them as noncentrosymmetric, is that we can use an incoming linearly
polarized beam of light
at normal incidence, and use the  direction of the polarized  electric
field to control $\calv^{\rma\rmb}(\go)$.
Indeed, writing 
$\bfE(\omega) = E_0(\omega)(\cos\ga\,\hat\bfx+\sin\ga\,\hat\bfy)$,
where $\alpha$ is the polarization angle, we obtain from
 Eq. \eqref{eq:vab-w}
that
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= 
\frac{2}{\hbar\xi(\go)}
\left(\mu^{\mathrm{abxx}}(\omega)\cos^{2}\alpha + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}\alpha + 
\mu^{\mathrm{abxy}}(\omega)\sin 2\alpha\right)
,
\label{eq:vab-aw}
\end{align}
\end{widetext}
for the structures chosen in this article,
$\xi^{\mathrm{xx}}(\go)=\xi^{\mathrm{yy}}(\go)\equiv\xi(\go)$, and $\xi^{\mathrm{xy}}(\go)=0$.
Now, we formalize our two options for $\calv^{\rma\rmb}(\go)$.

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define the magnitude of
the electron velocity on the plane of the structure 
with the spin polarized along the $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\gs^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
% where this angle is measured in the counter-clockwise direction from the
% positive $x$. 
We also define two special angles
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular
to the incoming polarization, 
respectively. The subscript $\gs^\rmb$ denotes the spin along $\rmb$.

@@

Fixing the calculated velocity along $\rma=x$ or $\rma=y$,
we define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented
in the $xyz$ system's coordinates
according to a polar angle
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
= 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{align}
and an azimuthal angle
\begin{align}
\varphi_{\mathrm{a}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}

@@

We present the results of $\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ and
$\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ for the C$_{16}$H$_{8}$-up  and
C$_{16}$H$_{8}$-alt structures, being both noncentrosymmetric semi-infinite 2D
carbon systems with 50\% hydrogenation in different arrangements. We recall
that the Up structure has hydrogen atoms only on the upper side of the carbon
sheet, while the Alt structure has alternating hydrogen atoms on the upper and
bottom sides. Also, we take the hexagonal carbon lattice to be on the $xy$ plane
for both structures, and the carbon-hydrogen bonds on the perpendicular $xz$
plane, as depicted in Figs. \ref{fig:up-struc} and \ref{fig:alt-struc}. The
coordinates for the Up and Alt unit cells of the structures are presented in
Tables \ref{tab:up-unitcell} and \ref{tab:alt-unitcell}.

We calculated the self-consistent ground state and the Kohn-Sham states using
density functional theory in the local density approximation (DFT-LDA), with a
planewave basis using the ABINIT code \cite{gonzeCPC09}.
% 
We used Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space
Gaussian pseudopotentials \cite{hartwigsenPRB98}, including the spin-orbit
interaction needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented
in Eq. \eqref{eq:mu}.
% 
The convergence parameters for the calculations of our results corresponding to
the Up and Alt structures are cutoff energies of 40\,Ha and 65\,Ha, resulting
in LDA energy band gaps of 0.084\,eV and 0.718\,eV, respectively. The energy
eigenvalues and matrix elements for the Up and Alt structures were calculated
using 12802 $\mathbf{k}$ points and 14452 $\mathbf{k}$ points in the IBZ to
integrate $\mu^{\mathrm{abcd}}(\go)$ and $\xi^{\rma\rmb}(\go)$ using the
linearized analytic tetrahedron method (LATM).\cite{nastosPRB07} We neglect the
anomalous velocity term $\hbar(\bfgs\times\nabla V)/4m^2c^2$, where $V$ is the
crystal potential, in $\hat\bfv$ of Eq.~\eqref{z.1}, as this term is known to
give small contribution to PSC.\cite{bhatPRL05} Therefore,
$[\hat\bfv,\hat\bfS]=0$, and Eq.~\eqref{z.1} reduces to $\hat K^{\rma\rmb}=\hat
v^\rma \hat S^\rmb=\hat S^\rmb \hat v^\rma$. Finally, the prime in the sum of
Eq.~\eqref{eq:mu} is restricted to quasidegenerated conduction bands $c$ and
$c'$ that are closer than 30 meV, where this value is both a typical laser-
pulse energy width and the room-temperature energy.\cite{nastosPRB07}
% 
% It is known that the DFT with LDA approximation misestimates the energy band
% gap of semiconductors. To correct this one has to include the many-body
% interaction using i.e. the so-called GW approximation but this technique has a
% very high computational cost and is out of the scope in this paper.
% Nevertheless, DFT still remains a mainstream tool for calculating diverse
% properties derived from the electronic band structure.

@@

In this subsection, we calculate $\calv_{\gs^\rmz}(\omega,\alpha)$, Eq.
\eqref{eq:vs-mag}, for the case where the spin is fixed along $z$, i.e.
directed perpendicularly to the surface of the Up and Alt structures. Also, we
calculate $\gamma_{\gs^\rmz}(\go,\ga)$, Eq. \eqref{eq:gamma-ang}, which
determines the direction along which the injected electrons move along the
surface of  each structure. We mention that we have also made the analysis for
the cases when the spin  is directed along $\mathrm{x}$ or $\mathrm{y}$,
finding similar qualitative results to those presented below.
%

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{upplots/up-vx-vy-w1}
\caption{For the Up structure, we show
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$
(left scale, black line),  
$\theta_{\mathrm{a}} (\omega,\alpha)$
(right scale, red line), and
$\varphi_{\mathrm{a}} (\omega,\alpha)$,
(right scale, blue line),
vs. $\hbar\go$, for $\ga=35^\circ$, 
and $\rma=x,y$.
}
\label{fig:up-vab-comp-rtp-1}
\end{figure}
